Using Scipion for Stream Image Processing at Cryo-EM Facilities

### 2018-09-24 Review/answer_referee.tex

\documentclass[a4paper,12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{fontenc}
\usepackage{graphicx}
\usepackage{scrextend}
\usepackage{tikz}
\usepackage{pgfplotstable}%color table
\usepackage{xspace}
\usepackage{url}

\def\cnb{CNB\xspace}

\newcommand{\ffigure}[1]{Figure \ref{#1}}
\newcommand{\ttable}[1]{Table \ref{#1}}
\newcommand{\ffigures}{Figures}
\newcommand{\ttables}{Tables}
\def\scipionbox{\textit{ScipionBox}\xspace}
\def\mnas{\textit{microscope-NAS}\xspace}
\def\onas{\textit{Output-NAS}\xspace}
\def\hserver{\textit{HTML-Server}\xspace}
\def\scipion{\textit{Scipion}\xspace}
\def\emadmin{\textit{EMAdmin}\xspace}

\def\epu{\textit{EPU}\xspace}
\def\scilifelab{SciLifeLab\xspace}
\def\cnbcsic{CNB-CSIC\xspace}
\def\cnb{CNB\xspace}

@@

\begin{reviewer}
\section*{Reviewer 1}
\point{The graphical abstract is likely too detailed 
and would not show well when reduced to the image size that the final version has.}

\reply  Both reviewers complain that the graphical abstract need to be redesigned. The first reviewer complains that is ``...too detailed...'' while the second one claims that ``...is oversimplified...''. Since the graphical abstract is not compulsory  and both reviewers dislike it for completely opposite and irreconcilable reasons we have decided to drop it.

\point{The manuscript would benefit from a figure that outlines the layout of a typical Scipion system, showing computers, data connections, data flows in workflow scenarios, and how streaming is now implemented. At least for one of the three examples described, a figure would be helpful and make the manuscript more appealing to readers. }

\reply The following figure has been included in the manuscript.

\begin{figure}
  \centering
      \includegraphics[width=0.75\textwidth]{images/diagram.pdf}
  \caption{Graphical representation of the data flow through the CNB facility. Green, blue and red lines represent the microscope private network, the \cnb local area network and the public network respectively.}
  \label{fig:cnbpipeline}

\end{figure}
\clearpage

%\point{The English language at a few locations would benefit from further proof reading, but these are minor problems. }

%\reply tut-tuc-tuc

@@

\point{Minor corrections in pages: Abstract, 4, 6, 9, 10, 11, 13, 16, 17}

\reply All grammar, spelling typos, rephrasing and other minor corrections suggested by the reviewer have been followed except one. Below we comment on the non followed suggestions (page 9 and 10) and the modifications introduced in page 16 since involve the rewriting of a paragraph.

Q \the\numexpr\value{pointcounter}-1\relax.1: Page 9: ``...Explain what EPU is...''. EPU has already been introduced by the following two paragraphs.

\begin{quote}
 Many software packages have been developed to support automated data collection, such as 
 .., EPU (Thermo Fisher Scientific, 2018), ..., etc.
\end{quote}

\begin{quote}
...we comment on an additional software (called  \emadmin) that connects \scipion with the microscope acquisition software (\epu) and records the facility activity.
\end{quote}

Q \the\numexpr\value{pointcounter}-1\relax.1: Page 10: ``Please explain, what ciphers are''. Following reviewer 2 recommendations this section has been simplified and the sentence containing the word ``ciphers'' has been deleted.

Q \the\numexpr\value{pointcounter}-1\relax.2: Page 16: ``The sentence -They can potentially play with the pipeline in ways that are not possible when having a programmer writing the code for them.- Is not understandable on first reading.''. We have rewriting the paragraph as:

\begin{quote}
 Workflows implement pipelines and describe them as a set of interconnected steps. In a workflow approach, users (including microscope staff) can reasonably be expected to modify the paths and incorporate new steps using a GUI. Alternate solutions as scripts do not clearly separate each step definition from the connections between the different steps and, for no programmers, are scary and difficult to modify. 
\end{quote}

@@

\end{reviewer}

\begin{reviewer}
\section*{Reviewer 2}

\point{The authors claim that the strength of their software is in its ability to process data while being collected; a point that is not well-addresses by other software. At the same time, they acknowledge that UCSFImage4, Focus, and Relion are software packages similar to Scipion capable of processing data in a high-throughput manner while new data is being acquired at the microscope. A clear differentiation between Scipion and these software packages was lacking in this manuscript. What advantages Scipion provide over the above mentioned software packages?}

[[[[[[[[[[[[[[[[[[[THIS ANSWER NEED TO BE POLISHED]]]]]]]]]]]]]]]]]]]]]]

\reply Since there is already a published manuscript describing \scipion, in the present manuscript  we decided to focus on \scipion features that have been specifically developed for working at EM facilities (e.g. streaming). The main reasons why \scipion is better suited than other packages for EM work is rooted in \scipion main design rather than the new facilities related features. Nevertheless, the reviewer is right in pointing out than a ``A clear differentiation between Scipion and these software packages was lacking in this manuscript''. Therefore, we have decided to add the following paragraphs to the manuscript

\begin{quote}

At the software architectural level,  the greatest difference 
between \scipion and the above mentioned software is that  \scipion pays 
special attention to define abstraction layers to simplify maintenance
and extensibility, and that follows a strict workflow approach. A few examples will illustrate the main advantages.

To create a new protocol in Scipion,
only a Python script needs to be developed. From this script, the
system will discover the new protocol and will automatically build
a form and store the protocol related information in a database.
\scipion protocol developers must certainly know \scipion conventions
but they do not need to modify or even understand the existing code.
In  \scipion next version, the abstraction will be increased
since protocols will be implemented as plug-ins,
that is, independent small python scripts  that add functionality to the main
\scipion application. The plug-in approach allows parallel development. That is,  
since protocols are implemented independently from the main application
they can be developed and released in parallel by different teams. In this way, support for the new version of \textit{Relion} does not require a new \scipion release but
a release of the independent \textit{Relion} plug-in

\scipion follows a pure workflows approach and therefore implement pipelines and describe them as a set of interconnected steps (or protocols in \scipion). In a workflow approach, users (including microscope staff) can reasonably be expected to modify the paths and incorporate new steps using a GUI. The previously mentioned software packages do not clearly separate each step definition from the connections between the different steps and, for no programmers, create or modify workflows may be challenging.

Another advantage of a clear separation between steps and relationships 
is the improvement of load valance, reproducibility and  provenance (process of tracing and recording the origins of data and its movement between protocols). The first step in a \scipion workflow execution
is to store in a database each of the protocols to be executed as well as how 
outputs of one protocol are attached to the input of another protocol. Then, a workflow engine analyzes the stored data and using as constrain the number of CPUs assigned by the user to the workflow optimizes the execution, that is, try to execute as many protocols in parallel as possible. Obviously, the data stored in the database  have enough information to execute each step and therefore guaranties reproducibility and provenance.
\end{quote}

%In Appion, however, the developer must develop the script, the cor-
%responding form, and an extra table in the database; 
%Appion developers might thus need to know the underlying framework in more
%detail.

%It is difficult to predict how EM software will evolve in the
%future. Our view is that software developers will continue to add
%algorithms to the different EM packages, but that the burden of
%many operations will be shifted from packages to frameworks.
%Bookkeeping will require special attention to provide real tracking
%and reproducibility. Workflows will have a key role when explor-
%ing processing alternatives. Most algorithms will need to use dis-
%tributed computing through clusters or the Cloud. Scipion is our
%first step in implementing an integrative framework to address
%important problems in the field simply and effectively

@@

\point{The authors mentioned that Scipion has been installed and in production mode in seven facilities. Is there a way to know how Scipion is being used and to get statistical data on its usage?}

\reply The following paragraphs have been added in the manuscript.

\begin{quote}

In the context of free and open source projects it is difficult to create usage statistics that are accurate. But certainly, if we want \scipion to succeed, we need to 
know how frequent the different protocols are used. This is the reason why we have recently developed our \textit{Scipion Usage Data Collector} that monitors -if activated by the user- the usage of the different protocols and send -via HTTP protocol- the information to the developer's team. This information is anonymous and cannot be used to identify the original computer in which \scipion has been executed.

Using the HTTP header it is possible to guess the country in which \scipion has been installed. With these information we have created the usage map and table available at URL \url{http://scipion.i2pc.es/report_protocols/scipionUsage/} and  \url{http://scipion.i2pc.es/report_protocols/protocolTable/}. The data corresponds on ``global'' \scipion usage rather than ``facility'' usage but there 
is a one to one relationship between the  countries with highest number of \scipion projects and the countries with \scipion based EM facilities. The only exception to the rule is Canada where two of the original \scipion developers are setting up a new facility (McGill University).

Note that, in a effort to report real use and skip test projects, the number of projects in the map refers to the number of non empty \scipion projects updated more than once. That is, 
\scipion projects that have been opened in at least two different days. Another constraint is that developers computers are filtered out.
\end{quote}

@@

\point{Section 4 on Scipion setup at different facilities was confusing and unnecessarily lengthy. It wasn't clear to me if the "Additional software developments" made in 4.1.2 and 4.2.2 were general to Scipion or only made as a case-by-case scenario?}

\reply Section 4 has been extensively rewritten to take into account the reviewers' suggestions. In summary, we have included a new image, delete several paragraphs, modify others and move those paragraphs related with future work to a different section.

Please confer the file ``Revised Manuscript(marked)'' attached to this review in which 
all significant differences between the original manuscript and the revised one are marked up.

Q \the\numexpr\value{pointcounter}-1\relax.1: ``It wasn't clear to me if the Additional software developments made in 4.1.2 and 4.2.2 were general to Scipion or only made as a case-by-case scenario?''. The following paragraph have been modified.

\begin{quote}
 This section in divided in two parts. In the first one we discuss the network setup that allows us to move the data from the microscope to the user laboratory. Then,  we comment on an additional software called  \emadmin. \emadmin has been specifically designed for the \cnb facility, it connects \scipion with the microscope acquisition software (\epu) and records the facility activity.
\end{quote}

@@

\point{The graphical abstract is oversimplified and does not convey any of the useful information presented in the manuscript. The background is distracting, the cartoons are redundant with the labels, and the image describing the reporting feature is not legible and improperly sized (laterally stretched).}

\reply See answer to reviewer 1 question 1.

\point{Finally, the manuscript could benefit greatly from more visual summaries and schematics to guide the readers especially the non-experts that the authors refer to as part of their target consumer base.}

\reply We think the reviewer refer here to section 4. As commented previously this section has been extensively rewritten to take into account the reviewers' suggestions and it includes a new figure.

### 2018-09-24 Review/0.abstract.tex
to home

Three dimensional electron microscopy is becoming a very data-intensive field in which vast amounts of experimental images are acquired at high speed. To manage such large-scale projects, we had previously developed  a modular workflow system called \scipion \citep{delarosatrevin2016}. We present here a major extension of \scipion that allows stream processing of EM images while the data is being acquired. This approach helps to detect problems at early stages, saves computing time and provides users with a detailed evaluation of the data quality before the acquisition is finished. %Different EM facilities have 
%implemented their stream processing solution using custom-made scripts, but this script based approach is not flexible and present a slow response time to new feature requests. The advantages of our proposed solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or with a different data set; (3) monitorization: there are utilities that constantly check how the execution of the algorithms is going on and inform about possible anomalies; (4) reports: we have developed several graphical interfaces that are updated periodically producing a summary (e.g, CTF defocus values, system load, etc). This summary may be generated in HTML format and  published through a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided. 
At present, \scipion has been installed and is in production mode in seven Cryo-EM facilities throughout the world, and it is being considered in many more.

>>>>

Three dimensional electron microscopy is becoming a very data-intensive field in which vast amounts of experimental images are acquired at high speed. To manage such large-scale projects, we had previously developed  a modular workflow system called \scipion \citep{delarosatrevin2016}. We present here a major extension of \scipion that allows processing of EM images while the data is being acquired. This approach helps to detect problems at early stages, saves computing time and provides users with a detailed evaluation of the data quality before the acquisition is finished. %Different EM facilities have 
%implemented their stream processing solution using custom-made scripts, but this script based approach is not flexible and present a slow response time to new feature requests. The advantages of our proposed solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or with a different data set; (3) monitorization: there are utilities that constantly check how the execution of the algorithms is going on and inform about possible anomalies; (4) reports: we have developed several graphical interfaces that are updated periodically producing a summary (e.g, CTF defocus values, system load, etc). This summary may be generated in HTML format and  published through a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided. 
At present, \scipion has been deployed and is in production mode in seven Cryo-EM facilities throughout the world.%, and it is being considered in many more.

### 2018-09-24 Review/2.overall.tex
to home

Stream processing, i.e. computing on data directly as they are produced or received, is the main change introduced in \scipion. 
In this way, a new level of parallelization has been reached since the  data generated by the microscope continuously flows through a chain of protocols. For example, as movies are imported to the system they are aligned and their CTF is estimated. In \scipion version 1.2, users may find  the first steps in the image processing workflow adapted to streaming, including movie alignment,  CTF estimation and particle picking and extraction. In the developers version, several 2D classification algorithms are being adapted for this new way of processing.

@@

\scipion is a flexible framework that allows to create different workflows by choosing among many algorithms at each step. Although the best workflow for each case will depend on specific requirements of specimens, microscopes or facilities, we  described the pipeline offered by default at the \emph{National Center for Biotechnology} (CNB) cryoEM Facility as use case to clarify Scipion capabilities.

>>>>

Stream processing, i.e., computing on data directly as they are produced or received, is the main change introduced in \scipion. 
In this way, a new level of parallelization has been reached since the  data generated by the microscope continuously flows through a chain of protocols. For example, as movies are imported to the system they are aligned and their CTF is estimated. In \scipion version 1.2, users may find  the first steps in the image processing workflow adapted to streaming, including movie alignment,  CTF estimation and particle picking and extraction. In the developers version, several 2D classification algorithms are being adapted for this new way of processing.

@@

\scipion is a flexible framework that allows to create different workflows by choosing among many algorithms at each step. Although the best workflow for each case will depend on specific requirements of specimens, microscopes or facilities, we  describe the pipeline offered by default at the \emph{National Center for Biotechnology} (\cnb) cryoEM Facility as use case to clarify \scipion capabilities.

### 2018-09-24 Review/3.workflow.tex
version2
to home

The data flow created by an electron microscope requires a careful design of the IT infrastructures.  Different facilities have adopted different solutions and therefore, how \scipion  access data and interchange information may require minor adjustments. In this section we discuss  the  solutions applied by the different facilities in which \scipion is running.

\scipion is so far being used routinely in eight Cryo-EM facilities:
\begin{itemize}
 \itemsep0em

@@

\end{itemize}
and it is being considered in many more.

\scipion original design satisfied CNB requirements, in which \scipion automatically fetches newly recorded movies from a network mounted disk. After this first installation, \scipion was deployed on SciLifeLab (a centralized facility located at Stockholm), where a rigorous booking system is followed. There, \scipion was adapted to interface with this booking platform to produce better reports. Lately,  \scipion has been installed in  large synchrotrons  such as ESRF and Diamond,  where projects are handled by ISPyB (a customized laboratory information management system -LIMS- \citep{Delageniere2011:ispb}) and data are saved on a distributed file system. % that heavily penalizes access to lists of files. 
In this environment, \scipion needs to constantly interchange information with ISPyB and with other applications through a message queue system called \emph{ActiveQ}.% and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created.

In the following, through three use cases, we describe in detail how \scipion has been adapted to the different environments.

>>>>

The data flow created by an electron microscope requires a careful design of the IT infrastructures.  Different facilities have adopted different solutions and therefore, how \scipion  access data and interchange information may require minor adjustments. In this section we discuss  the  solutions applied by some facilities in which \scipion is running.

So far \scipion is being used routinely in seven Cryo-EM facilities:
\begin{itemize}
 \itemsep0em

@@

\end{itemize}
%and it is being considered in many more.

\scipion original design satisfied CNB requirements, in which \scipion automatically fetches newly recorded movies from a network mounted disk. After this first installation, \scipion was deployed on SciLifeLab (a centralized facility located at Stockholm), where a rigorous booking system is followed. There, middleware (software that acts as a bridge between applications) was developed to interface with this booking platform to produce better reports. Lately,  \scipion has been installed in  large synchrotrons  such as ESRF and Diamond,  where projects are handled by ISPyB (a customized laboratory information management system -LIMS- \citep{Delageniere2011:ispb}) and data are saved on a distributed file system. % that heavily penalizes access to lists of files. 
In this environment, \scipion needs to constantly interchange information with ISPyB and with other applications through a message queue system called \emph{ActiveQ}.% and it cannot longer check if there is a new file in a particular storage device but must wait to be informed by the system that a new file has been created.

%In the following, through three use cases, we describe in detail how \scipion has been adapted to the different environments.

In the following, through the three above mentioned use cases, we describe in detail how \scipion is been used in these different environments. %Notice that EM facilities are designed and implemented to support high-throughput data acquisition and require complex data-collection workflows.  These workflows need access to experimental parameters such as microscope settings, sample and user information, as well as channels for feeding processing results back to the control system.  
For each use case we will describe first the facility setup and then in the subsection entitled ``Aditional Software Developments'' we will comment on specific  software developed for each facility that connects \scipion with the facility and microscope control software.

### 2018-09-24 Review/3.workflow_cnb.tex
to home

\subsection{Case 1: Scipion at CNB}
The National Center for Biotechnology (\cnb) forms part of the Spanish National Research Council (CSIC), the largest public research institution in Spain. 
The \cnb Cryo-Electron Microscopy Facility is a joint effort of the Nacional Center for Biotechnology and the Biological Research Centers. The mission of the facility is to provide researchers access to the latest cryo-EM technology for high resolution imaging.

This section in divided in two parts. In the first one we discuss the network setup that allows us to move the data from the microscope to the user laboratory. Then  we comment on an additional software (called  \emadmin) that connects \scipion with the microscope acquisition software (\epu) and records the facility activity.

@@

At the beginning \cnb networking setup was a quite straightforward one, movies were stored in the same machine that ran \scipion and processed data were saved to external USB hard disks by this same machine. The situation changed when the new Falcon-III detector was installed. After that, data were produced at sustained speeds as high as 90 MB/sec (when acquiring four movies by hole in linear mode) and the main problem became to move smoothly the data through the entire pipeline from the microscope to the final user laboratory. In the following we describe such a pipeline.

From a network point of view there are three different levels of accessibility (see  \ffigure{fig:cnbpipeline}): (1) a private network that cannot be accessed by regular users -green lines in \ffigure{fig:cnbpipeline}-, (2) the \cnb LAN (local area network)  that can be accessed by any user which works inside the hosting  institution -blue lines in \ffigure{fig:cnbpipeline}- and (3) finally, there are two computers that may be accessed from outside the institution -red lines in \ffigure{fig:cnbpipeline}-, these computers give access to data and a web site respectively. s

\begin{figure}
  \centering
      \includegraphics[width=0.75\textwidth]{images/diagram.pdf}
  \caption{Graphical representation of the data flow through the CNB facility. Green, blue and red lines represent the microscope private network, the \cnb local area network and the public network respectively}
  \label{fig:cnbpipeline}

@@

Data sets are recorded using \epu. The microscope control system has a 70 TB NAS (Network-attached storage) unit for saving movies as they are obtained (we will refer to this NAS as microscope-NAS). This storage unit is shared through a Samba share and connected to the private network using a 10Gb connexion. \scipion is executed on a Linux server which has two 1Gb NICs (network interface controllers) that connect it to the private and the LAN networks respectively (we will refer to this machine as \scipionbox). \scipionbox has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 2 \textit{Quadro M4000} GPUs. Finally, a second NAS with four 1Gb NICs is linked to the private LAN and WAN networks (we will refer to this second NAS as output-NAS).

The data flow is as follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox reads the movies from the microscope-NAS but does not store them locally, (c) output-NAS has access to Microscope-NAS and \scipionbox hard disks (d) output-NAS has a collection of scripts than may be used to incrementally backup \scipion project and the movies to a local or remote disk, or serve this information through the Internet and, (e) in parallel, periodically \scipionbox transfers to  HTML Server reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).

Although CNB staff try to discourage the use of external USB hard disks versus network as way of retrieving the raw and processed data, this is the preferred option for most users.
USB 3.0 drives can theoretically handle data flows of 300 MB/s, but CNB experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s. In most cases, 50 MB/s is enough to transfer the data before the following microscope  turn-shift starts. When two, or even three, users try to download their
data simultaneously, output-NAS works as front-end and isolates the data
acquiring and processing infrastructure from the user's downloading activities.
Network users access to output-NAS using a hardened common account. This account only allows rsync connections and make impossible to retrieve data if the name of the project to be downloaded is unknown. Weak ciphers (as arcfour) has been activated in outputNAS to speed up transfer.

@@

\emadmin is free software available at github repository (\url{https://github.com/rmarabini/webservices/tree/master/EMadmin}). You are welcomed to download and customize it to meet your organization specific requirements.

>>>>

\subsection{Case 1: Scipion at CNB}
The National Center for Biotechnology (\cnb) forms part of the Spanish National Research Council (CSIC), the largest public research institution in Spain. 
The \cnb Cryo-Electron Microscopy Facility is a joint effort of the Nacional Center for Biotechnology and the Biological Research Centers. %The mission of the facility is to provide researchers access to the latest cryo-EM technology for high resolution imaging.

This section in divided in two parts. In the first one we discuss the network setup that allows us to move the data from the microscope to the user laboratory. Then,  we comment on an additional software called  \emadmin. \emadmin has been specifically designed for the \cnb facility, it connects \scipion with the microscope acquisition software (\epu) and records the facility activity.

@@

At the beginning \cnb networking setup was a quite straightforward one, movies were stored in the same machine that ran \scipion and processed data were saved to external USB hard disks by this same machine. The situation changed when the new Falcon-III detector was installed. After that, data were produced at sustained speeds as high as 90 MB/sec (when acquiring four movies by hole in linear mode) and the main problem became to move smoothly the data through the entire pipeline from the microscope to the final user laboratory. In the following we describe such a pipeline (see \ffigure{fig:cnbpipeline} for details.).

\begin{figure}
  \centering
      \includegraphics[width=0.75\textwidth]{images/diagram.pdf}
  \caption{Graphical representation of the data flow through the CNB facility. Green, blue and red lines represent the microscope private network, the \cnb local area network and the public network respectively.}
  \label{fig:cnbpipeline}

@@

%From a network point of view there are three different levels of accessibility (see  \ffigure{fig:cnbpipeline}): (1) a private network that cannot be accessed by regular users -green lines in \ffigure{fig:cnbpipeline}-, (2) the \cnb LAN (local area network)  that can be accessed by any user who works inside the hosting  institution -blue lines in \ffigure{fig:cnbpipeline}- and (3) finally, there are two computers that may be accessed from outside the institution -red lines in \ffigure{fig:cnbpipeline}-, these computers give access to data and a web site respectively.

Data sets are recorded using \epu. The microscope control system has a 70 TB NAS (Network-attached storage) unit for saving movies as they are obtained (we will refer to this NAS as \mnas). This storage unit is shared through a Samba share and connected to the private network using a 10Gb connexion. \scipion is executed on a Linux server which has two 1Gb NICs (network interface controllers) that connect it to the private network and the LAN (local area network) respectively (we will refer to this machine as \scipionbox). \scipionbox has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 2 \textit{Quadro M4000} GPUs. Finally, a second NAS with four 1Gb NICs is linked to the private, LAN and WAN networks (we will refer to this second NAS as \onas) so it can retrieve data from any computer and serve it to external users.

The data flow is as follows: (a) the microscope system stores movies in the \mnas, (b) when executing movie-alignment protocols \scipionbox reads the movies from the \mnas but does not store them locally, (c) \onas has access to \mnas and \scipionbox hard disks (d) \onas has a collection of scripts than may be used to incrementally backup \scipion project and the movies to a local or remote disk, or serve this information through the Internet and, (e) in parallel, periodically \scipionbox transfers to  \hserver reports that may be checked by users or staff with an HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).

Although \cnb staff try to discourage the use of external USB hard disks versus network as way of retrieving the raw and processed data, this is the preferred option for most users.
USB 3.0 drives can theoretically handle data flows of 300 MB/s, but \cnb experience is that, in this busy environment, many of them  have trouble exceeding 50 MB/s. In most cases, 50 MB/s is enough to transfer the data before the following microscope  turn-shift starts. %When two, or even three, users try to download their
%data simultaneously, \onas works as front-end and isolates the data
%acquiring and processing infrastructure from the user's downloading activities.
Network users access to \onas using a hardened common account. This account only allows rsync connections and make impossible to retrieve data if the name of the project to be downloaded is unknown. %The achievable speed of copying a file depends on quite a few different factors 
%including the selected ciphers (algorithms for performing data encryption or decryption). Weak ciphers (as arcfour) have been activated in \onas to speed up transfer.

@@

\emadmin is free software available at github repository (\url{https://github.com/rmarabini/webservices/tree/master/EMadmin}). The software can be freely downloaded and customized to meet the specific requirements of client organizations.

### 2018-09-24 Review/3.workflow_diamond.tex
to home

Diamond Light Source is the UK National Synchrotron User Facility.  As such, it provides access to large-scale experimental facilities to the UK and worldwide scientific communities to conduct experiments that are not possible in their home laboratories.  As a consequence of this national role, Diamond has significant experience at managing many visiting scientists and user experiments, and the associated data and post-processing.  %In the area of Macromolecular Crystallography, this has become a highly automated process which allows users to visualise electron density maps within minutes of the experiment being completed (REF. Aller P et al Methods Mol Biol 2015)
%The recent ``Revolution Resolution'' (REF. Kulbrandt W, Science 2014) in cryoEM has transformed the impact and demand of cryoEM in Structural Biology. However, the instrumentation and running costs of the latest generation of EM necessary for high-end cryoEM is beyond scope of all but the best-funded home laboratories. Due to the huge demand for high-end microscope access, 
Recently, the electron Bio-Imaging Centre (eBIC) has been set up at Diamond~\citep{diamond2017}. The centre houses four Thermofisher Titan Krios, a Talos Arctica and a Scios dual-beam. All of the transmission electron microscopes are equipped with counting-mode direct electron detectors and Volta phase plates.

@@

Diamond has thousands of users every year, both academic and commercial, and it is vital that their data is kept secure and confidential. A user management system
has been developed to allocate each user with a unique federal identity (ID),
and to associate this ID with the experiments in which the user partakes in.
 Visits are structured such that all raw and processed data collected during an experiment are kept in one directory, and only the associated visitors have read and write permissions in that directory. Users may access their data through SynchWeb which fronts ISPyB, a Laboratory Information Management System combining sample tracking and experiment reporting \citep{Delageniere2011:ispb, Fisher2015:synchweb}.

@@

Although most of developments described in this subsection will be integrated in the main \scipion code repository, at present they are available in a developer's branch  at \url{https://github.com/I2PC/scipion/tree/release-1.1-headless-devel}.

>>>>

Diamond Light Source is the UK National Synchrotron User Facility.  As such, it provides access to large-scale experimental facilities to the UK and worldwide scientific communities to conduct experiments that are not possible in their home laboratories.  As a consequence of this national role, Diamond has significant experience at managing many visiting scientists and user experiments, and the associated data and post-processing.  %In the area of Macromolecular Crystallography, this has become a highly automated process which allows users to visualise electron density maps within minutes of the experiment being completed (REF. Aller P et al Methods Mol Biol 2015)
%The recent ``Revolution Resolution'' (REF. Kulbrandt W, Science 2014) in cryoEM has transformed the impact and demand of cryoEM in Structural Biology. However, the instrumentation and running costs of the latest generation of EM necessary for high-end cryoEM is beyond scope of all but the best-funded home laboratories. Due to the huge demand for high-end microscope access, 
Recently, the electron Bio-Imaging Centre (eBIC) has been set up at Diamond~\citep{diamond2017}. The centre houses four Thermo Fischer Scientific Titan Krios, a Talos Arctica and a Scios dual-beam. All of the transmission electron microscopes are equipped with counting-mode direct electron detectors and Volta phase plates.

@@

Diamond has thousands of users every year, both academic and commercial, and it is vital that their data is kept secure and confidential. A user management system
has been developed to allocate each user with a unique federal identity (ID),
and to associate this ID with the experiments in which the user partakes in.
 %Visits are structured such that all raw and processed data collected during an experiment are kept in one directory, and only the associated visitors have read and write permissions in that directory. 
 Users may access their data through SynchWeb which fronts ISPyB, a Laboratory Information Management System combining sample tracking and experiment reporting \citep{Delageniere2011:ispb, Fisher2015:synchweb}.

@@

At present, most of developments described in this subsection are available in a developer's branch  at \url{https://github.com/I2PC/scipion/tree/release-1.1-headless-devel}. Some of them (e.g. the monitor protocol) will be integrated in the main \scipion repository while the more Diamond specific software will be published in an independent repository. [** Ask Pablo about headless]

### 2018-09-24 Review/3.workflow_scilifelab.tex
to home

The Swedish National Cryo-EM Facility offers access to state-of-the-art equipment and expertise in single particle cryo-EM and cryo electron tomography (cryo-ET). The Facility has two nodes: at \scilifelab in Stockholm and at Ume\aa\ University. \scilifelab in Stockholm offers single-particle service with a 
Talos Arctica for sample optimisation and a Titan Krios for high-resolution data collection. The Ume\aa\ node is expected to become operational during 2018, and will offer cryo-ET with a Titan Krios and a Scios DualBeam SEM.

The facilities can be accessed by Swedish researchers through 
a peer-reviewed process. Applications can be submitted through an application portal (\url{https://cryoem.scilifelab.se/}). % and will be evaluated once every three months based on their scientific merit and technical feasibility by a national Project Evaluation Committee. %On the other side, some time of the facility is reserved for internal research groups at Stockholm University. 
Internally, an online booking system called \emph{Booked Scheduler} centralizes the reservations of the microscopes and other instruments.

@@

To avoid interruption of data collection due to network issues each microscope has its own dedicated storage, that in the case of the Titan microscope is double since it may be connected to two different cameras, a K2 and a Falcon III. Hence a script to continuously move the produced data from the adequate storage to the staging server has been created.

@@

\subsubsection{Additional Software Developments: \emph{Session Wizard}}
In order to interface \scipion with the microscope infrastructure an upgraded version of the original script developed at CNB was used. Unfortunately, it was difficult to reuse the existing code due to a different folder structure and, more important, a different approach to user and microscope management. %To address the new set of requirements, a new wizard is under development at \scilifelab. For this new wizard, a basic data model was designed representing the current processing and data organization at \scilifelab. %The model is composed by the following entities:

@@

When a user starts a data collection, the wizard is executed using the command \textit{session-titan} (or \textit{session-talos}). The wizard determines which user has booked the microscope by fetching today's booking information from \emph{Booked Scheduler} through their web API. From the reservation information the wizard will also detect if it is an internal or a national facility project. 
In case of the latter, additional information will be retrieved from the Application Portal (such as principal investigator, project code, contact person or invoice address). After gathering all this information the wizard stores it, creates the processing environment and launches \scipion. %For example, the movies-import protocol will already point to the location where the files will be written for this session. Moreover, taking into account which microscope-camera is being  used, the files pattern can be inferred.

The current implementation of the wizard stores each session in a simple database file. This database, together with extra information from the Booking System and the Application Portal, is used to generate invoices and reports for a given period. %The generation of a session report is under development, that will provide a summary and useful information for users to access the acquired data. 
%[** this last sentence is not very informative]

The described session wizard can be found at  \url{https://github.com/delarosatrevin/scipion-session}. You are welcomed to download and customize it to meet your organization specific requirements. \emph{Booked Scheduler} is available at (\url{https://www.bookedscheduler.com/})

>>>>

The Swedish National Cryo-EM Facility %%offers access to state-of-the-art equipment and expertise in single particle cryo-EM and cryo electron tomography (cryo-ET). The Facility 
has two nodes: at \scilifelab in Stockholm and at Ume\aa\ University. \scilifelab in Stockholm offers single-particle service with a 
Talos Arctica for sample optimisation and a Titan Krios for high-resolution data collection. The Ume\aa\ node is expected to become operational during 2018, and will offer cryo-ET with a Titan Krios and a Scios DualBeam SEM.
%The facilities can be accessed by Swedish researchers through 
%a peer-reviewed process. Applications can be submitted through an application portal (\url{https://cryoem.scilifelab.se/}). % and will be evaluated once every three months based on their scientific merit and technical feasibility by a national Project Evaluation Committee. %On the other side, some time of the facility is reserved for internal research groups at Stockholm University. 
Internally, an online booking system called \emph{Booked Scheduler} centralizes the reservations of the microscopes and other instruments.

@@

%To avoid interruption of data collection due to network issues each microscope has its own dedicated storage, that in the case of the Titan microscope is double since it may be connected to two different cameras, a K2 and a Falcon III. Hence a script to continuously move the produced data from the adequate storage to the staging server has been created.

@@

\subsubsection{Additional Software Developments: \emph{Session Wizard}}
In order to interface \scipion with the microscope infrastructure an upgraded version of the original script developed at CNB was used. Unfortunately, it was difficult to reuse the existing code due to a different folder structure and, more important, a different approach to user and microscope management. To address the new set of requirements, a new wizard has been development at \scilifelab. (In this contest wizard refers to a helper script which guide users through a certain process.) %For this new wizard, a basic data model was designed representing the current processing and data organization at \scilifelab. %The model is composed by the following entities:

@@

When a user starts a data collection, the wizard is executed using the command \textit{session-titan} (or \textit{session-talos}). The wizard determines which user has booked the microscope by fetching today's booking information from \emph{Booked Scheduler} through their web API, %From the reservation information the wizard will also detect if it is an internal or a national facility project. 
%In case of the latter, additional information will be retrieved from the Application Portal (such as principal investigator, project code, contact person or invoice address). 
%After gathering all this information the wizard 
stores it, creates the processing environment and launches \scipion. %For example, the movies-import protocol will already point to the location where the files will be written for this session. Moreover, taking into account which microscope-camera is being  used, the files pattern can be inferred. 
The current implementation of the wizard stores each session in a simple database file. This database, together with extra information from the Booking System and an Application Portal, is used to generate invoices and reports for a given period. %The generation of a session report is under development, that will provide a summary and useful information for users to access the acquired data. 
%[** this last sentence is not very informative]

The described session wizard can be found at  
\url{https://github.com/delarosatrevin/scipion-session}. The software can be freely downloaded and customized to meet the specific requirements of client organizations. \emph{Booked Scheduler} is available at (\url{https://www.bookedscheduler.com/})
